---
title: "We Talk R 15 mins/day"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General approach

We will keep a tally of these activities here.  Every day, 15 minutes of R activity.

# Project based

## Project 1:  Trust as a latent entity

Let's figure out how to model Trust as a latent entity.  What you will learn:

1. EFA
2. Model inspection
3. Fit evaluation
4. Psychometrics
5. R coding

Things to do:
1. Decide on the data
2. Get some objectives line up
3. Operationalize said objectives

### Data

We are using the 5th study data (SCM dissertation).

```{r Data}
dat <- read.csv("./dat5alll.csv",header=TRUE)[,-1]
library(psych)
describe(dat)
```

### Objectives

Create an EFA model for trust from the available data.  Use the factor scores of the first factor to predict self-reported trust (T) that ought to, in turn predict behavioral intent (B) to trust the agent.

#### EFA

```{r EFA}
## We'll use G, U1, U2, and R as manifest indicators of F1 (latent entity of
## trust)

library(psych)
efa1 <- fa(dat[,3:6],missing=TRUE) 
efa1
## compute the correlation matrix to check relationships bw vars
corPlot(cor(dat[,3:7],use="pairwise.complete.obs"))
## compute the factor scores for F1 and add to dat
dat$efaF1 <- factor.scores(dat[,3:6],efa1)[[1]]

```

```{r Models}

```

```{r EvalModels}

```

You can include R code in the document as follows:

```{r cars}
summary(cars)
```
